### 2012-11-16 Krantz-v3/Chapters/chapter1/HSLDA.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

@@

HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We
call individual groups of bag-of-word data documents (although they could be any data represented by bags of features).  Let $w_{n,d}
\in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
= \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document
$d$.  Let there be $D$ such documents and let the size of the vocabulary be
$V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label
$l \in \mathcal{L}$, except the root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
also in the set of labels. 
 We will for exposition purposes assume that this label set has hard ``is-a''
 parent-child constraints (explained later), although this assumption can be
 relaxed at the cost of more computationally complex inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we demonstrate, only positive labels are observed.   This may not be true of all applications; however, positive-only label imbalance is a common problem.  How we solve this problem will be discussed later.

The constraints imposed by an is-a label hierarchy are that if the $l$th label is applied to document $d$, i.e.,~$y_{l,d} = 1$, then all labels in the label hierarchy up to the root are also applied to document $d$, i.e.,~$y_{\mathrm{pa}(l),d} = 1, y_{\mathrm{pa}(\mathrm{pa}(l)),d} = 1, \ldots, y_{r,d}=1.$  Conversely, if a label $l'$ is marked as not applying to a document then no descendant of that label may be applied to the same.   We assume that at least one label is applied to every document.  This is illustrated in Figure~\ref{fig:graphical_model} where the root label is always applied but only some of the descendant labelings are observed as having been applied (diagonal hashing indicates that potentially some of the plated variables are observed).
%To reiterate, these is-a label hierarchy constraints can be relaxed.
%To capture this hierarchical structure we model the labeling of documents
%using a generative cascade of conditional probit regression models.

@@

In HSLDA, the bag-of-word document data is modeled using LDA with
 full, hierarchical topic estimation (i.e. with the global topic estimated too rather than assumed).
%In HSLDA, documents are modeled using the LDA mixed-membership mixture model
%with global topic estimation. 
Label responses are modeled using a conditional
hierarchy of probit regressors. The HSLDA graphical model is given in
Figure~\ref{fig:graphical_model}.

 In the corresponding mathematical generative description of HSLDA to follow, $K$ is the number of LDA
``topics'' (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$
is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific
distribution over topics, $\boldsymbol\beta$ is a global distribution over
topics, Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution,
$\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution,
$\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the
$d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator
function that takes the value $1$ if its argument is true and $0$ otherwise.

\subsection{Ancestral sampling for HSLDA}

The following procedure describes how to generate from the HSLDA generative
model.

@@

%We will now describe the stochastic generative process which defines
%our model. The 
%Given the number of topics, K, and broad gamma priors on hyperparameters,
%the generative process is as follows: 
\begin{shortbox}
\Boxhead{HSLDA Ancestral Sampler}
\begin{enumerate}
\item For each topic $k=1,\ldots,K$

@@

\end{itemize}
\end{itemize}
\end{enumerate}
\end{shortbox}

@@

The second half of step 4 is what is referred to as supervision in the supervised LDA literature.  This is where the hierarchical classification of the bag-of-words data takes place and the is-a label constraints are enforced.  For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.  Note that $y_{l,d}$ can only be applied to document $d$  (set to 1) if its parent label $\mathrm{pa}(l)$ was also applied (these expressions are specific to is-a constraints but could be modified to accommodate different constraints between labels).  The regression coefficients $\boldsymbol\eta_l$ are independent a priori, however, the hierarchical coupling in this model induces a posteriori dependence.   The net effect of this conditional hierarchy of profit regressors is that label predictors deeper in the label hierarchy are able to focus on finding features that distinguish the members of that set, conditioned on the fact that the members are in the same set.  It should be clear that one can trivially restrict this hierarchy to a depth one hierarchy.  Also, one can nearly as easily make the conditional classification at each node multi-class rather than single-class if more than one label at each node is required.  In many cases, however, a binary indicator along with a deeper or more complex tree is sufficient.
%We believe this to be a significant source of the empirical label prediction improvement we observe experimentally.  We test this hypothesis in Section~\ref{sec:experiments}.

Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations.  In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:regression_param_conditional} which are analytic.  This simplifies posterior inference substantially.  A review of probit regression can be found near the end of this chapter in Section \ref{sec:probit_review}.

@@

In the common case where no negative labels are observed (like the example applications we consider in Section~\ref{sec:experiments}), the model must be explicitly biased towards generating  negative labels in order to keep it from learning to assign positive labels to all documents.   This is a common problem in modeling unbalanced data.  To see how this model can achieve this we draw the reader's attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.  Because $\bar{\mathbf{z}}_d$ is always positive, setting $\mu$ to a negative value results in a bias towards negative labelings, i.e.~for large negative values of $\mu$, all labels become a priori more likely to be negative ($y_{l,d}=-1$).  We explore the effect of $\mu$ on out-of-sample label prediction performance in Section~\ref{sec:experiments}.   In a very real way $\mu$ is a knob that can be adjusted both before inference to induce a broad array of out-of-sample performance characteristics that vary along classical axes like specificity, recall, and accuracy.  A similar but less principled solution can be effected by changing the decision boundary from $0$ in \eqref{zero_0} and \eqref{zero_1}.  This technique can be used to vary out-of-sample label bias after learning.

>>>>

% !TEX root =  ../../Krantz_style.tex

@@

This model (HSLDA) is designed to fit hierarchically, multiply-labeled, bag-of-word data.  We
call groups of bag-of-words data documents (unordered words in text documents, bag of visual feature representations of images, etc.).  Let $w_{n,d}
\in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
= \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document
$d$.  Let there be $D$ such documents and let the size of the vocabulary be
$V=|\Sigma|$.

Let the set of labels be $\mathcal{L}=\left\{
  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.  A label in HSLDA corresponds to a node in the graphical model in Figure~\ref{fig:graphical_model}.  A label can either be observed or unobserved.  Documents can be multiply labeled, meaning that subsets of label nodes in the graphical model in Fig.~\ref{fig:graphical_model} can have observed values.  Each label
$l \in \mathcal{L}$, except the root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
also in the set of labels. 
 We will for exposition purposes assume that this label set has hard ``is-a''
 parent-child constraints (explained later), although this assumption can be
 relaxed at the cost of more computationally complex inference.  Such a label hierarchy forms a multiply rooted tree.   Confused readers may wish to consult Figure \ref{fig:diagnosis_tree} or Figure \ref{fig:product_tree} each of which is a  label-tree graphical model.

 In a label forest a node may be observed (the label was ``applied'' and, for instance, was observed to have value 1), unobserved and unknown, or unobserved and constrained to be either -1 or 1 by the structure of the label space.   Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label applies to document $d$ or not.   In most cases $y_{l,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we demonstrate, only positive labels are observed.   This may not be true of all applications; however, positive-only label imbalance is a common problem.  How we solve this problem will be discussed later.

The constraints imposed by an is-a label hierarchy are that if the $l$th label is applied to document $d$, i.e.,~$y_{l,d} = 1$, then all labels in the label hierarchy up to the root are also applied to document $d$, i.e.,~$y_{\mathrm{pa}(l),d} = 1, y_{\mathrm{pa}(\mathrm{pa}(l)),d} = 1, \ldots, y_{r,d}=1.$  Conversely, if a label $l'$ is marked as not applying to a document (i.e.~$y_{l',d}=-1$) then no descendant label of that label can take value 1.   We assume that at least one label is applied to every document.  This is illustrated in Figure~\ref{fig:graphical_model} where the root label is always applied but only some of the descendant labelings are observed as having been applied (diagonal hashing indicates that potentially some of the plated variables are observed).
%To reiterate, these is-a label hierarchy constraints can be relaxed.
%To capture this hierarchical structure we model the labeling of documents
%using a generative cascade of conditional probit regression models.

@@

In HSLDA, the bag-of-word document data is modeled using LDA with
 full, hierarchical topic estimation (i.e. with the global topic estimated too rather than assumed).
%In HSLDA, documents are modeled using the LDA mixed-membership mixture model
%with global topic estimation. 
Label responses are modeled using a conditional
hierarchy of probit regressors and will be discussed next. The full HSLDA graphical model is given in
Figure~\ref{fig:graphical_model}.

\subsection{Generative Model}

In the following box  the HSLDA generative model  is given for the ``is-a hierarchy'' set of label constraints.   In the box and what follows in this chapter, $K$ is the number of LDA
``topics'' (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$
is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific
distribution over topics, $\boldsymbol\beta$ is a global distribution over
topics, Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution,
$\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution,
$\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the
$d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator
function that takes the value $1$ if its argument is true and $0$ otherwise.

@@

%We will now describe the stochastic generative process which defines
%our model. The 
%Given the number of topics, K, and broad gamma priors on hyperparameters,
%the generative process is as follows: 
\begin{shortbox}
\Boxhead{HSLDA Generative Model}
\begin{enumerate}
\item For each topic $k=1,\ldots,K$

@@

\end{itemize}
\end{itemize}
\end{enumerate}
\label{box:generative_model}
\end{shortbox}

@@

The second half of step 4 is what is referred to as supervision in the supervised LDA literature.  This is where the hierarchical classification of the bag-of-words data takes place and the is-a label constraints are enforced.  For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.    \eqref{zero_0} and \eqref{zero_1} comprise a probit regression model in an auxiliary variable formulation (see Section \ref{sec:probit_review}).   Note that in the case that the parent label is applied, i.e.~$y_{\mathrm{pa}(l),d}=1,$ the child label $y_{l,d}$ is applied with probability $P(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l}>0).$  This is a conditional probit regression model for classification where $\boldsymbol\eta_l$ are the class-conditional regression parameters.  The auxiliary variables $a_{l,d}$ make inference tractable but are not fundamental to the model -- only the labels and regression parameters are actually of interest.

Note that $y_{l,d}$ can only be applied to document $d$  (set to 1) if its parent label $\mathrm{pa}(l)$ was also applied (these expressions are specific to is-a constraints but could be modified to accommodate different constraints between labels).  Note that multiple labels can be applied to the same document.  The regression coefficients $\boldsymbol\eta_l$ which generate the labels are independent a priori, however, the hierarchical coupling in this model and conditional label dependency structure induces a posteriori dependence.   The net effect of this conditional hierarchy of profit regressors is that child label predictors deeper in the label hierarchy are able to focus on finding features that distinguish label paths in the tree, conditioned on the fact that all the children of any particular node are by design members of some more general parent set.  It should be clear that one can trivially restrict this hierarchy to a depth one hierarchy recovering SLDA with probit link and univariate categorical labels.  Also, one can nearly as easily make the conditional classification at each node multi-class rather than single-class if more than one label at each node is required.  In many cases, however, a binary indicator along with a deeper or more complex tree is sufficient.
%We believe this to be a significant source of the empirical label prediction improvement we observe experimentally.  We test this hypothesis in Section~\ref{sec:experiments}.

Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations (see Appendix \ref{sec:probit_review}).  In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:regression_param_conditional} which are analytic.  This simplifies posterior inference substantially.  A review of probit regression can be found near the end of this chapter in Section \ref{sec:probit_review}.

@@

\subsection{Dealing with Label Imbalance}

In the common case where no negative labels are observed (like the example applications we consider in Section~\ref{sec:example_applications}), the model must be explicitly biased towards generating  negative labels in order to keep it from learning to only assign positive labels to all documents.   This is a common problem in modeling with unbalanced labels.  To see how this model can achieve this we draw the reader's attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.  Because $\bar{\mathbf{z}}_d$ is always positive, setting $\mu$ to a negative value results in a bias towards negative labelings, i.e.~for large negative values of $\mu$, all labels become a priori more likely to be negative ($y_{l,d}=-1$).  We explore the effect of $\mu$ on out-of-sample label prediction performance in Section~\ref{sec:example_applications}.   In a very real way $\mu$ is a knob that can be adjusted both before inference to induce a broad array of out-of-sample performance characteristics that vary along classical axes like specificity, recall, and accuracy.  A similar but less principled solution can be effected by changing the decision boundary from $0$ in \eqref{zero_0} and \eqref{zero_1}.  This technique can be used to vary out-of-sample label bias after learning.

### 2012-11-16 Krantz-v3/Chapters/chapter1/discussion.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

@@

The SLDA model family, of which HSLDA is a member, can be understood in two
different ways. One way is to see it as a family of topic models that improve
on the topic modeling performance of LDA via the inclusion of observed
supervision. An alternative, complementary way is to see it as a set of models
that can predict labels for bag-of-word data. A large diversity of problems
can be expressed as label prediction problems for bag-of-word data. A
surprisingly large amount of that kind of data possess structured labels,
either hierarchically constrained or otherwise.  HSLDA directly addresses
this kind of data and works well in practice. That it
outperforms more straightforward approaches should be of interest to
practitioners.
%
%Variational Bayes has been the predominant estimation approach applied to sLDA
%models. Hierarchical probit regression makes for tractable Markov chain Monte
%Carlo SLDA inference, a benefit that should extend to other sLDA models should
%probit regression be used for response variable prediction there too.
%
%The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
%it is better to do full joint estimation of HSLDA.  An alternative
%interpretation of the same results is that, if one is more sensitive to the
%performance gains that result from exploiting the structure of the labels, then
%one can, in an engineering sense, get nearly as much gain in label prediction
%performance by first fitting LDA and then fitting a hierarchical probit
%regression.  There are applied settings in which this could be advantageous.

>>>>

% !TEX root =  ../../Krantz_style.tex

\section{Related Work}
\label{sec:related_work}

HSLDA does not, of course, stand alone.  Models for structured labeling of bag of words data can be designed in a number of different ways.

As shown  in Section~\ref{sec:example_applications},  SLDA can be used to solve this kind of problem directly; however, doing so requires ignoring the hierarchical
dependencies amongst the labels.  Other models that incorporate LDA and supervision and so also could be used to solve this problem include
LabeledLDA~\cite{Ramage2009} and DiscLDA~\cite{DiscLDA}.  Various applications of these models to 
computer vision and document networks have been explored~\cite{wangbleifeifei08,RelationalLDA}.
None of these models, however, leverage dependency structure in the label space.

In other non-LDA-based related work, researchers have classified documents into hierarchies (a closely related task) using naive Bayes classifiers
and support vector machines. Most of this work has been demonstrated on relatively
small datasets, small label spaces, and has focused on single label classification without
a model of documents such as LDA~\cite{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.  
\section{Discussion}
\label{sec:discussion}

@@

The SLDA model family, of which HSLDA is a member, can be understood in two
different ways. One way is to see it as a family of topic models that improve
on the topic modeling performance of LDA via the inclusion of observed
supervision. An alternative, complementary way is to see it as a set of models
that can predict labels for bag-of-word data. A large diversity of problems
can be expressed as label prediction problems for bag-of-word data. A
surprisingly large amount data possess structured labels,
either hierarchically constrained or otherwise.  HSLDA directly addresses
this kind of data and works well in practice. That it
outperforms more straightforward approaches should be of interest to
practitioners.
%
%Variational Bayes has been the predominant estimation approach applied to sLDA
%models. Hierarchical probit regression makes for tractable Markov chain Monte
%Carlo SLDA inference, a benefit that should extend to other sLDA models should
%probit regression be used for response variable prediction there too.
%
%The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
%it is better to do full joint estimation of HSLDA.  An alternative
%interpretation of the same results is that, if one is more sensitive to the
%performance gains that result from exploiting the structure of the labels, then
%one can, in an engineering sense, get nearly as much gain in label prediction
%performance by first fitting LDA and then fitting a hierarchical probit
%regression.  There are applied settings in which this could be advantageous.

There are many kinds of problems that have the same characteristics as this: any data  that consists of free text that has been partially or
completely categorized by human editors for instance; more fully
% In this paper we focus on
any bag-of-words data that has been, at least in part, 
categorized. 
 Examples include but are not limited to webpages and curated
hierarchical directories of the same \cite{DMOZ}, product descriptions and
catalogs, (e.g.~\cite{AMAZON} as available from \cite{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes.
The model we cover in this chapter shows one way to combine 
these two sources of information into a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. %The models and techniques that we develop in
%this paper are applicable to other data as well, namely, any unstructured
%representations of data that have been hierarchically classified (e.g.,~image
%catalogs with bag-of-feature representations).

%There are several challenges entailed in incorporating a hierarchy of labels
%into the model. Among them, given a large set of potential labels (often thousands), each
%instance has only a small number of labels associated to it. Furthermore, there are no
%naturally occurring negative labeling in the data, and the absence of a label
%cannot always be interpreted as a negative labeling.

%Our work operates within the framework of topic modeling. Our approach learns
%topic models of the underlying data and labeling strategies in a joint model,
%while leveraging the hierarchical structure of the labels. 
%In this chapter we focus on ``is-a'' hierarchically structured label spaces, but label spaces that are structured differently may be accommodated with relatively minor modifications.

### 2012-11-16 Krantz-v3/Chapters/chapter1/example_applications.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

Having described HSLDA and given details on how to do inference in it, we turn to demonstrations of using HSLDA to solve problems from  two different domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions.

@@

We compare HSLDA to two closely related models. The comparison models are SLDA with independent
regressors (hierarchical constraints on labels ignored,  i.e. the regression is not conditional) and HSLDA fit by first
performing LDA then fitting tree-conditional regressions (rather than jointly inferring the topics and the regression coefficients). These models were
chosen because they are the strongest available competitors and because they  highlight several pedagogical aspects of HSLDA including performance in the
absence  of hierarchical constraints, the effect of the combined inference, and
regression performance attributable solely to the hierarchical constraints.

@@

The number of topics for all models was set to $K=50$, the prior distributions of
$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
$p\left(\gamma\right)$ were all chosen to be gamma with a shape parameter of 1 and a
scale parameter of 1000.

@@

These results indicate that the full HSLDA model predicts more of the the
correct labels at a cost of an increase in the number of false positives
relative to the comparison models.

>>>>

% !TEX root =  ../../Krantz_style.tex

%Having described HSLDA and given details on how to do inference in it, we turn to demonstrations of using HSLDA to solve problems from  two different domains: predicting medical
%diagnosis codes from hospital discharge summaries and predicting product
%categories from Amazon.com product descriptions.

@@

We compare HSLDA to two closely related models. The comparison models are SLDA with independent
regressors (hierarchical constraints on labels ignored,  i.e. the regression is not conditional) and HSLDA fit by first
performing LDA then fitting probit regressors that respect the conditional label hierarchy (rather than jointly inferring the topics and the regression coefficients). These models were
chosen because they are the strongest available competitors and because they  highlight several pedagogical aspects of HSLDA including performance in the
absence  of hierarchical constraints, the effect of the combined inference, and
regression performance attributable solely to the hierarchical constraints.

@@

The number of topics for all models was set to $K=50$, the prior distributions of
$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
$p\left(\gamma\right)$ were all chosen to be gamma with a shape parameter of 1 and a
scale parameter of 1000.   Different values of $K$ corresponding to different numbers of topics were explored, however, the results that we show in the following are not substantially changed in character.  As is usual in mixed-membership models, there is an ideal number of topics that should be used for out-of-sample prediction tasks, however, a full model-selection search varying topic cardinality was not performed for these datasets.

@@

These results indicate that the full HSLDA model predicts more of the the
correct labels at a cost of an increase in the number of false positives
relative to the comparison models.  However, as shown in Figure~\ref{fig:1a}  HSLDA outperforms or no worse than the comparison models across across the full range of specificities.
